Energy and the CES production function

### 2015-10-20 Articles/Article_EcolEcon_2015/EcolEcon_2015A.Rnw
updates using "fake" objects of classes cdeModel and cesModel

<<Models_CSP>>=
fileNameCSP <- "EcolEcon2015ModelsCSP.rds"
if (!rerunModels){
  # load from the file on disk
  modelsCSP <- readRDS(fileNameCSP)
} else {
  #
  # Models that adhere to the cost-share theorem
  #

  modelsCSP <- list()
  # Output elasticities based on cost share principle
  alphaCSP <- 0.3 # Fraction of national accounts spent on capital
  betaCSP <- 0.7 # Fraction of national accounts spent on labor
  for (country in countries){
    for (adj in c("unadjusted", "adjusted")){
      # Replace with better code later. 
      # Would like to say cdModel(formula, data, factorShares=c(alpha_1 = 0.3, alpha_2 = 0.7, alpha_3 = 0))
      # or something like that.
      form <- keep_summands(formulas[[country]][[adj]], n=-3) # Remove energy
      yVar <- as.name(summands(form, n=1, right=FALSE))
      kVar <- as.name(summands(form, n=1))
      lVar <- as.name(summands(form, n=2))
      tVar <- as.name(summands(form, n=3))
      data <- historicalData[[country]]
      # Cobb-Douglas
      cdModCSP <- 
        lm(log(eval(yVar)) - alphaCSP*log(eval(kVar)) - betaCSP*log(eval(lVar)) ~ eval(tVar), 
           data = data)
      # cdModCSP is an "lm" object. It is not a "CDEmodel" object. 
      # Thus, it needs its own "naturalCoeffs" attribute if it is to work correctly in the "naturalCoef" function.
      attr(cdModCSP, "naturalCoeffs") <- data.frame(lambda=coef(cdModCSP)[[2]],
                                                    logscale=coef(cdModCSP)[[1]], 
                                                    scale=exp(coef(cdModCSP)[[1]]),
                                                    alpha_1=alphaCSP,
                                                    alpha_2=betaCSP, 
                                                    alpha_3=0
                                                    )
      cdModCSP$formula <- form
      sdata <- subset(data, select=all.vars(form))
      sdata <- data[complete.cases(sdata), unique(c(all.vars(form), names(data)))]
      cdModCSP$data <- sdata
      cdModCSP$response <- sdata$iGDP
      modelsCSP[[country]][[adj]][["noE"]][["CD"]] <- cdModCSP
      # CES
      # The only CES models that can adhere to the CSP are models where
      # alpha values are constant with respect to time. 
      # I.e., models where energy is not included AND where rho_1 = 0.
      # CES models where rho_1 = 0 are the Cobb-Douglas model.
      # Convienently, we previously fit that model.
      cesModCSP <- cdModCSP
      # Again, we need a customized "naturalCoeffs" attribute so that 
      # the "naturalCoef" function will do the right thing.
      # "naturalCoeffs" can be based on the same "naturalCoeffs"
      # from the Cobb-Douglas model, but a few additions are needed.
      attr(cesModCSP, "naturalCoeffs") <- 
        naturalCoef(cdModCSP) %>%
        mutate(
          delta_1 = naturalCoef(cdModCSP)[["alpha_1"]],
          delta = 1,
          rho_1 = 0, 
          rho = NA,
          sigma_1 = 1,
          sigma_12 = 1,
          sigma = NA,
          sigma_12_3 = NA
        )
      # Add output elasticities (as a function of time) to this model.
      # This would have been done by a "fortify" function, 
      # but this is not a typical CES model, so we can't "fortify" it.
      # Add other metadata, too.
      cesModCSP$data <- cesModCSP$data %>%
        mutate(
          oe1 = alphaCSP,
          oe2 = betaCSP,
          oe3 = 0,
          Country = country, 
          flavor  = ifelse(adj=="unadjusted", "Unadjusted", "Quality-adjusted"),
          energy  = "Without energy",
          model   = "CES", 
          nest    = "kl",
          CSP     = "w/ CSP"
        )
      modelsCSP[[country]][[adj]][["noE"]][["CES"]][["kl"]] <- cesModCSP
    }
  }
  # Save cost-share theorem models to disk.
  saveRDS(modelsCSP, file=fileNameCSP)
}

>>>>

<<Models_CSP>>=
fileNameCSP <- "EcolEcon2015ModelsCSP.rds"
if (FALSE && !rerunModels){
  # load from the file on disk
  modelsCSP <- readRDS(fileNameCSP)
} else {
  #
  # Models that adhere to the cost-share theorem
  #

  modelsCSP <- list()
  # Output elasticities based on cost share principle
  alphaCSP <- 0.3 # Fraction of national accounts spent on capital
  betaCSP <- 0.7 # Fraction of national accounts spent on labor
  for (country in countries){
    for (adj in c("unadjusted", "adjusted")){
      print(adj)
      # Replace with better code later. 
      # Would like to say cdModel(formula, data, factorShares=c(alpha_1 = 0.3, alpha_2 = 0.7, alpha_3 = 0))
      # or something like that.
      form <- keep_summands(formulas[[country]][[adj]], n=-3) # Remove energy
      yVar <- as.name(summands(form, n=1, right=FALSE))
      kVar <- as.name(summands(form, n=1))
      lVar <- as.name(summands(form, n=2))
      tVar <- as.name(summands(form, n=3))
      data <- historicalData[[country]]
      formula <- substitute(
        log(Y) - alpha_1 * log(K) - alpha_2 * log(L) ~ T,
        list( Y = yVar, K = kVar, L = lVar, T = tVar, A = alphaCSP, B = betaCSP)
      )
      # Cobb-Douglas
      cdModCSP <- 
        eval(substitute( 
          plm(formula, data = data, params = c(alpha_1 = 0.3, alpha_2 = 0.7, alpha_3 = 0.0), 
              optimize = FALSE),
          list(formula = formula)))

      # convert hand-fit object into a CDEmodel object
      names(cdModCSP$coefficients)[1:2] <- c("logscale", "lambda")
      class(cdModCSP) <- c("CDEmodel", class(cdModCSP))
      cdModCSP$formula <- form
      sdata <- subset(data, select=all.vars(form))
      sdata <- data[complete.cases(sdata), unique(c(all.vars(form), names(data)))]
      cdModCSP$data <- sdata
      cdModCSP$response <- sdata$iGDP
      cdModCSP$handcrafted <- TRUE

      # store the result in our list
      modelsCSP[[country]][[adj]][["noE"]][["CD"]] <- cdModCSP
      # CES
      # The only CES models that can adhere to the CSP are models where
      # alpha values are constant with respect to time. 
      # I.e., models where energy is not included AND where rho_1 = 0.
      # CES models where rho_1 = 0 are the Cobb-Douglas model.
      # Convienently, we previously fit that model.
      cesModCSP <- cdModCSP

      cesModCSP$coefficients <- 
        c(cesModCSP$coefficients, 
          c(delta_1 = cesModCSP$coefficients["alpha_1"],
            delta = 0,
            rho_1 = 0,
            rho = NA)
        )
      cesModCSP$handcrafted <- TRUE
      class(cesModCSP) <- setdiff(c("cesModel", class(cesModCSP)), "CDEmodel")
      # # Again, we need a customized "naturalCoeffs" attribute so that 
      # # the "naturalCoef" function will do the right thing.
      # # "naturalCoeffs" can be based on the same "naturalCoeffs"
      # # from the Cobb-Douglas model, but a few additions are needed.
      # attr(cesModCSP, "naturalCoeffs") <- 
      #   naturalCoef(cdModCSP) %>%
      #   mutate(
      #     delta_1 = naturalCoef(cdModCSP)[["alpha_1"]],
      #     delta = 1,
      #     rho_1 = 0, 
      #     rho = NA,
      #     sigma_1 = 1,
      #     sigma_12 = 1,
      #     sigma = NA,
      #     sigma_12_3 = NA
      #   )
      # # Add output elasticities (as a function of time) to this model.
      # # This would have been done by a "fortify" function, 
      # # but this is not a typical CES model, so we can't "fortify" it.
      # # Add other metadata, too.
      # cesModCSP$data <- cesModCSP$data %>%
      #   mutate(
      #     oe1 = alphaCSP,
      #     oe2 = betaCSP,
      #     oe3 = 0,
      #     Country = country, 
      #     flavor  = ifelse(adj=="unadjusted", "Unadjusted", "Quality-adjusted"),
      #     energy  = "Without energy",
      #     model   = "CES", 
      #     nest    = "kl",
      #     CSP     = "w/ CSP"
      #   )
      modelsCSP[[country]][[adj]][["noE"]][["CES"]][["kl"]] <- cesModCSP
    }
  }
  # Save cost-share theorem models to disk.
  saveRDS(modelsCSP, file=fileNameCSP)
}
